\section{Model Equations}

Mass, momentum and energy transport equations are expressed for
The gas phase and the $N$ solid phases. 
These equations are coupled through constitutive equations expressing
the interphase exchange terms. Closure equations expressing the
physical properties and the nature of the interpenetrating fluids are 
expressed to obtain a consistent equation system.

%----------------------------------------------------------------------
\subsection{Mass Conservation Equations.}
%----------------------------------------------------------------------
The continuity equations express the transport of gas and solid bulk
density. No mass transfer is allowed between different phases, no
chemical reactions between gas component are described. 

\subsubsection{\hspace{1cm} Mass conservation of the gas phase}

\begin{equation}
\frac{\partial}{\partial t}(\epsilon_{g}\rho_{g})+
\mbox{\boldmath $\nabla$}\cdot (\epsilon_{g}\rho_{g}{\bf v}_{g})=0,
\label{eq:gas_mass}
\end{equation}

%----------------------------------------------------------------------

\subsubsection{\hspace{1cm}Mass conservation of $N$ solid phases}

\begin{equation}
\frac{\partial}{\partial t}(\epsilon_{k}\rho_{k})+
\mbox{\boldmath $\nabla$}\cdot (\epsilon_{k}\rho_{k}{\bf v}_{k})=0,
\qquad k=1,2, ....,N.
\label{eq:solid_mass}
\end{equation}

%----------------------------------------------------------------------

\subsubsection{\hspace{1cm}Transport of gas chemical components}

\begin{equation}
\frac{\partial}{\partial t}(\epsilon_{g}\rho_{g}y_{i})+
\mbox{\boldmath $\nabla$}\cdot (\epsilon_{g}\rho_{g}y_{i}{\bf v}_{g})=0,
\qquad i=1,2,...,M.
\label{eq:gc_mass}
\end{equation}

%----------------------------------------------------------------------
\subsection{Momentum Transport Equations.}
%----------------------------------------------------------------------
The transient momentum transport equations for gas and particles are
similar and account for convective and 
diffusive (viscous) transport, pressure and gravity force. Gas and solids
equations are coupled through the gas-particle drag term. Different particles
interaction is described by an analogous particle-particle drag term.

\subsubsection{\hspace{1cm}Momentum balance for the gas phase}

\begin{equation}
\frac{\partial}{\partial t}(\epsilon_{g}\rho_{g}{\bf v}_{g})+
\mbox{\boldmath $\nabla$}\cdot (\epsilon_{g}\rho_{g}{\bf v}_{g}{\bf v}_{g})=
-\epsilon_{g}\mbox{\boldmath $\nabla$}P_{g}+
\mbox{\boldmath $\nabla$}{\bf T}_{g}
+\epsilon_{g}\rho_{g}{\bf g}+\sum_{k=1}^{N} D_{g,k}({\bf v}_{k}-{\bf v}_{g})
\label{eq:gas_mom}
\end{equation}

%----------------------------------------------------------------------

\subsubsection{\hspace{1cm}Momentum balance for $N$ solid phases}

\begin{eqnarray}
\frac{\partial}{\partial t}(\epsilon_{k}\rho_{k}{\bf v}_{k})+
\mbox{\boldmath $\nabla$}\cdot (\epsilon_{k}\rho_{k}{\bf v}_{k}{\bf v}_{k}) =
-\epsilon_{k}\mbox{\boldmath $\nabla$}P_{g}+
\mbox{\boldmath $\nabla$}{\bf T}_{k}
+\epsilon_{k}\rho_{k}{\bf g} -
D_{g,k}({\bf v}_{k}-{\bf v}_{g}) + 
\sum_{j=1}^{N} D_{k,j}({\bf v}_{j} - {\bf v}_{k}) \\
\label{eq:solid_mom}
k,j=1,2,...,N; \quad j \neq k
\nonumber
\end{eqnarray}

%----------------------------------------------------------------------
\subsection{Enthalpy Transport Equations.}
%----------------------------------------------------------------------
Energy trasport equations are expressed in terms of thermodynamic enthalpy.
The form of gas and solids equations are slightly different, because
particle-particle heat exchange is neglected. Viscous dissipation heat 
release is also neglected. Diffusive transport is expressed in the form
of Fourier law. Gas and particle equations are coupled through
the heat exchange term.

\subsubsection{\hspace{1cm}Enthalpy balance for the gas phase}

\begin{equation}
\frac{\partial}{\partial t}(\epsilon_{g}\rho_{g}h_{g})+
\mbox{\boldmath $\nabla$}\cdot (\epsilon_{g}\rho_{g}h_{g}{\bf v}_{g})=
\epsilon_{g}\bigg(\frac{\partial P_{g}}{\partial t}+{\bf v}_{g}\cdot
\mbox{\boldmath $\nabla$}P_{g}\bigg) +
\mbox{\boldmath $\nabla$}\cdot (k_{ge}\epsilon_{g}
\mbox{\boldmath $\nabla$}T_{g})+\sum_{k=1}^{N} Q_{k}(T_{k}-T_{g})
\label{eq:gas_enth}
\end{equation}

%----------------------------------------------------------------------

\subsubsection{\hspace{1cm}Enthalpy balance for $N$ solid phases}
\begin{equation}
\frac{\partial}{\partial t}(\epsilon_{k}\rho_{k}h_{k})+
\mbox{\boldmath $\nabla$}\cdot (\epsilon_{k}\rho_{k}h_{k}{\bf v}_{k})=
\mbox{\boldmath $\nabla$}\cdot (k_{ke}\epsilon_{k}\mbox{\boldmath 
$\nabla$}T_{k})
-Q_{k}(T_{k}-T_{g}), \hbox{\qquad $k = 1,2,...,N$}
\label{eq:solid_enth}
\end{equation}

%----------------------------------------------------------------------
\subsection{Constitutive and Closure Equations}
%----------------------------------------------------------------------
Closure relations the fluid density in terms of thermodynamic pressure 
and temperature, and the stress tensor in terms of the strain tensor.
The relation between temperature and enthalpy is also given.
A perfect gas equation of state and Newtonian (viscous) stress tensor
for gas are assumed. Smagorinsky closure expresses the turbulent 
dissipation in terms of an effective turbulent viscosity.
For particles a correction (Coulombic repulsion) to the Newtonian 
approximation accounts for particle finite size. Volumetric and mass
fraction closure express the fact that phases are treated as 
interpenetrating continua.

\subsubsection{\hspace{1cm} Volumetric and mass fraction closures}

\begin{equation}
\epsilon_{g} + \sum_{k=1}^{N}\epsilon_{k} = 1; 
\qquad \sum_{i=1}^{M}y _{i} = 1
\label{eq:closeps}
\end{equation}

%----------------------------------------------------------------------
\subsubsection{\hspace{1cm}Thermal equations of state}

\begin{equation}
\rho_{g}=\frac{P_{g}}{\tilde{R} T_g}; 
\hspace{10mm}\rho_{k}=\rho_{sk},\qquad k=1,2,...,N
\label{eq:th_eosg}
\end{equation}

%----------------------------------------------------------------------

\subsubsection{\hspace{1cm}Caloric equations of state}

\begin{equation}
T_{g}=\frac{h_{g}}{C_{p_{g}}}, 
\hspace{10mm} T_{k}=\frac{h_{k}}{C_{p_{k}}}, \qquad k=1,2,...,N
\label{eq:caloric_eos}
\end{equation}

\begin{equation}
C_{p_{g}}=\sum_{i=1}^{M}y _{i} C_{p_{g,i}}.
\label{eq:gas_cp}
\end{equation}
 
%----------------------------------------------------------------------
\subsubsection{\hspace{1cm}Gas-phase stress tensor}

\begin{equation} 
{\bf T}_{g}= \epsilon_{g}\Big( \mu_{g} \mbox{\boldmath $\widetilde{\tau}_{g}$} +
             \mu_{gt}\mbox{\boldmath ${\tau}_{g}$} \Big)
\label{eq:gas_stress}
\end{equation}

\begin{eqnarray}
\mbox{\boldmath ${\tau}_{g}$} & = & \Big[\mbox{\boldmath $\nabla$}
{\bf v}_{g}+ (\mbox{\boldmath $\nabla$}{\bf v}_{g})^{T}\Big] \\
tr(\mbox{\boldmath ${\tau}_{g}$}) & = & \frac{2}{3}(\mbox{\boldmath $\nabla$}\cdot{\bf v}_{g}){\bf I} \\
\mbox{\boldmath $\widetilde {\tau}_{g}$} & = & \mbox{\boldmath ${\tau}_{g}$} - tr(\mbox{\boldmath ${\tau}_{g}$}) 
\label{eq:strain}
\end{eqnarray}

\begin{eqnarray} 
\mu_{gt} & = & l^2 \rho_g |\mbox{\boldmath ${\tau}_{g}$}| \\ 
\label{eq:gas_mu}
l & = & l_S = c_S \Delta ; \quad \Delta = \sqrt{\Delta r \Delta z}\\
l & = & l_B = \kappa (z + z_0); 
\label{scale}
\end{eqnarray}

\begin{eqnarray}
k_{gt} & = & \frac{C_{pg}\mu_{gt}}{Pr_t} \\
k_{ge} & = & k_{g} + k_{gt}
\label{eq:prandtl}
\end{eqnarray}

%----------------------------------------------------------------------
\subsubsection{\hspace{1cm}Solid stress tensor}

\begin{equation}
{\bf T}_{k} = \epsilon_{k} \mu_{k} \mbox{\boldmath $\widetilde\tau$}_{v,k} - 
              \tau_{c,k} {\bf I},
\hbox{\qquad $k=1,2,...,N$}
\label{eq:solid_stress}
\end{equation}

\begin{eqnarray}
\mbox{\boldmath $\widetilde\tau$}_{v,k} & = & 
\Big[ \mbox{\boldmath $\nabla$}{\bf v}_{k}+
(\mbox{\boldmath $\nabla$}{\bf v}_{k})^{T} \Big]-
\frac{2}{3} (\mbox{\boldmath $\nabla$}\cdot {\bf v}_{k}){\bf I}
\label{eq:solid_viscous}\\
\mbox{\boldmath $\nabla$} \tau_{c,k} & = & 
G(\epsilon_k) \mbox{\boldmath $\nabla$} \epsilon_k,
\qquad G(\epsilon_k) = 10^{-a \epsilon_k + b}
\label{eq:solid_coulombic}
\end{eqnarray}

%----------------------------------------------------------------------
\subsection{Interphase Exchange Terms.}
%----------------------------------------------------------------------
Interphase exchange coefficients are derived from semi-empirical correlations.
Their validity can depend on the flow regime, thus including also 
correction for turbulence and different particle concentrations.

\subsubsection{\hspace{1cm}Gas-particle Drag Coefficient}

for $\epsilon_g \geq 0.8$:

\begin{equation}
D_{g,k}=\frac{3}{4}C_{d,k}\frac{\epsilon_{g}\epsilon_{k}\rho_{g}\mid 
{\bf v}_{g}- {\bf v}_{k}\mid}{d_{k}}\epsilon_{g}^{-2.7}, 
\qquad k=1,2,...,N
\label{eq:drag_less}
\end{equation}
%
\begin{eqnarray}
C_{d,k} & = & \frac{24}{Re_{k}}[1+0.15\, Re_{k}^{0.687}], 
\qquad Re_{k} < 1000 \\
& = & 0.44, \qquad Re_{k} \geq 1000,
\label{eq:drag_coeff1}
\end{eqnarray}
%
\noindent
for $\epsilon_g < 0.8$:

\begin{equation}
D_{g,k}=150 \frac{\epsilon_{k}^{2}\mu_{g}}{\epsilon_{g} d_{k}^{2}} +
1.75 \frac{\epsilon_{k}\rho_{g}\mid {\bf v}_{g}-
{\bf v}_{k}\mid}{d_{k}}, \qquad k=1,2,...,N
\label{eq:drag_more}
\end{equation}

\begin{equation}
Re_{k}=\frac{\epsilon_{g}\rho_{g}d_{k}\mid 
{\bf v}_{g}-{\bf v}_{k}\mid}{\mu_{g}}
\label{eq:reynolds}
\end{equation}

%----------------------------------------------------------------------
\subsubsection{\hspace{1cm}Particle-particle Drag Coefficient}

\begin{equation}
D_{k,j} = F_{kj} \alpha (1+e) \rho_k \epsilon_{k} \rho_j \epsilon_{j}
\frac{(d_{k}+d_{j})^{2}}
{(\rho_{k}d_{k}^{3}+ \rho_{j}d_{j}^{3})} \mid {\bf v}_{k}-{\bf v}_{j}\mid,
\hspace{10mm} k,j = 1,2,...,N; \quad j \neq k
\label{eq:ppdrag}
\end{equation}\\

\begin{equation}
F_{kj} = \frac {3 \epsilon_{kj}^{1/3} + (\epsilon_k + \epsilon_j)^{1/3}} 
{2(\epsilon_{kj}^{1/3}-(\epsilon_k+\epsilon_j)^{1/3})}
\label{eq:effe}
\end{equation}\\

\noindent
for $\displaystyle X_k \leq \frac{\Phi_k}{\Phi_k+(1-\Phi_k)\Phi_j}$:

\begin{eqnarray}
\epsilon_{kj}=\big[(\Phi_k-\Phi_j)+(1-a)(1-\Phi_k)\Phi_j\big]
\frac {\big[\Phi_k + (1-\Phi_j)\Phi_k \big]} {\Phi_k} X_k + \Phi_j
%\nonumber
\end{eqnarray}\\

\noindent
for $\displaystyle X_k > \frac{\Phi_k}{\Phi_k+(1-\Phi_k)\Phi_j}$:

\begin{eqnarray}
\epsilon_{kj}=(1-a)\big[\Phi_k + (1-\Phi_k)\Phi_j \big] (1-X_k) + \Phi_k
%\nonumber
\end{eqnarray}
%
\begin{eqnarray}
a=\sqrt{\frac{d_j}{d_k}}, \qquad (d_k \geq d_j), 
\qquad X_k = \frac{\epsilon_k}{\epsilon_k + \epsilon_j}
%\nonumber
\end{eqnarray}\\
%

\subsubsection{\hspace{1cm}Gas-particle heat transfer coefficient}

%
\begin{eqnarray}
Q_{k}=Nu_{k}\frac{6k_{g}\epsilon_{k}}{d_{k}^{2}},
\hspace{10mm} k = 1,2,...,N
\label{eq:heattr}
\end{eqnarray}
%
\begin{eqnarray}
Nu_{k} = (2 + 5 \epsilon_k^2)(1+0.7Re_{k}^{0.2}Pr^{1/3})+
(0.13+1.2\epsilon_k^2)Re_k^{0.7}Pr^{1/3}, 
\qquad Re_{k} \leq 10^5
\label{eq:nusselt}
\end{eqnarray}
%
\begin{eqnarray}
Re_{k}=\frac{\rho_{g}d_{k}\mid{\bf v}_{g}-{\bf v}_{k}\mid}{\mu_{g}},
\qquad Pr=\frac{C_{pg}\mu_{g}}{k_g}
\label{eq:reypr}
\end{eqnarray}
%

\clearpage
