%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                          %
%              (C) Copyright 2003  INGV and CINECA                         %
%                      All Rights Reserved                                 %
%								  	   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{makeidx}
\usepackage{psfig}

\makeindex

% define margins, etc
\topmargin	0.1in
\oddsidemargin	0in
\evensidemargin	0in
\textheight	8.80in
\textwidth	6.50in
\marginparsep	0.25cm
\headheight	0in
\headsep	0in
%\footskip	0.5in
%\footheight	0in%

% define macros
\input{pdac_macros}

\setcounter{secnumdepth}{3}
%% \setcounter{tocdepth}{5} %% very detailed table of contents
\setcounter{tocdepth}{4}

%
% the document itself
%

\begin{document}

% EXPLORIS heading
%\input{Exploris_head}

% initial pages for the guide - title, TOC, copyright
%\input{pdac_ug_begin}
\input{pdac_begindoc}

% Introduction
\input{pdac_intro}

% Input and output files
\newpage
\input{pdac_io}

% Input simulation parameters
% (INPUT)
\newpage
\input{pdac_input}

% Sample configuration files
\newpage
\input{pdac_sample}

% Description of how to run PDAC
\newpage
\input{pdac_runit}

% How to get and install PDAC
\newpage
\input{pdac_avail}

\newpage
\addtocontents{toc}{\contentsline {section}{References}{\thepage}}
\begin{thebibliography}{10}
\bibitem{Dobran_etal1993}
F.~Dobran, A.~Neri and G.~Macedonio, Numerical simulation of collapsing volcanic columns (1993), {\em J. Geophys. Res.} 98~(B3), 4231--4259. 
\bibitem{Neri_Macedonio_1996}
Neri,~A. and G.~Macedonio, Numerical simulation of collapsing volcanic columns with particles of two sizes (1996), {\em J. Geophys. Res.} 101~(B04), 8153--8174.
\bibitem{Neri_etal_2003}
Neri~A., T.~Esposti~Ongaro, G.~Macedonio, D.~Gidaspow. (2003),
Multiparticles simulation of collapsing volcanic columns and pyroclastic
flows, {\em J. Geoph. Res.} 108(B4), 2202, doi:10.1029/2001JB000508
\bibitem{EspostiOngaro_etal_2007a}
Esposti~Ongaro,~T., A.~Neri, C.~Cavazzoni, G.~Erbacci and M.V.~Salvetti (2007a),
A parallel multiphase flow code for the 3D simulation of volcanic explosive eruptions, {\em Parallel Computing 33}, 541-560, doi:10.1016/j.parco.2007.04.003
\bibitem{EspostiOngaro_etal_2007b}
Esposti~Ongaro~T., A.~Clarke, A.~Neri, B.~Voight, and C.~Widiwijayanti (2007b), Fluid dynamics of the 1997 Boxing Day volcanic blast on Montserrat, W.I., {\em J. Geophys. Res.}, 113, B03211, doi:10.1029/2006JB004898
\bibitem{EspostiOngaro_etal_2008}
Esposti~Ongaro~T., A.~Neri, G.~Menconi, M.~De'Michieli~Vitturi, P.~Marianelli, C.~Cavazzoni, G.~Erbacci and P.J.~Baxter, A transient 3D numerical simulation of column collapse and pyroclastic flow sceanrios at Vesuvius (2008), {\em J. Volcanol. Geotherm. Res.}, {\it submitted}
\end{thebibliography}

\newpage
\addtocontents{toc}{\contentsline {section}{Index}{\thepage}}
\printindex

% that's all, folks
\end{document}
