
\thispagestyle{empty}

\section*{PDAC (\PDACNAME )\\
Non-Transferable, Non-Commercial Use License}

\subsubsection*{\PDAC: \PDACNAME\ LICENSE AGREEMENT}

Upon execution of this Agreement by the party identiﬁed below (“Licensee”), The Istituto Nazionale di Geoﬁsica e Vulcanologia (“INGV”), on behalf of the \PDAC\ Development Team (“PDAC Team”), will provide \PDAC\ in Executable Code and/or Source Code form (“Software”) to Licensee, subject to the following terms and conditions. For purposes of this Agreement, Executable Code is the compiled code, which is ready to run on Licensee’s computer. Source code consists of a set of files which contain the actual program commands that are compiled to form the Executable Code. 

\begin{enumerate}
\item Copyright and nature of the license

The main authors of \PDAC\ are Augusto Neri, Giovanni Macedonio and Tomaso Esposti Ongaro (the PDAC Team). The intellectual property of this Software is owned by the authors at INGV, and all right, title and interest, including copyright, remain with INGV. INGV grants, and Licensee hereby accepts, a non-exclusive, non-transferable license to use the Software for academic, research and internal business purposes only e.g. not for commercial use (see Paragraph 5 below), without a fee. 
Because substantial funds in the development of \PDAC\ have been provided by the European Commission  and Ministero dell’Istruzione, dell’Universit\`a e della Ricerca (Italy), any use, distribution or sublicense of the Software is subject to the specific policies of these institutions.

\item Copies and modifications

Licensee agrees to reproduce the copyright notice and other proprietary markings on all copies of the Software. Licensee has no right to transfer or sublicense the Software to any unauthorized person or entity. However, Licensee does have the right to make complementary works that interoperate with \PDAC\ and to freely distribute such complementary works.
Licensee may modify the Software to make derivative works, for its own academic, research, and internal business purposes. Licensee’s distribution of any derivative work is also subject to the same restrictions on distribution and use limitations that are speciﬁed herein for INGV Software. Prior to any such distribution the Licensee shall require the recipient of the Licensee’s derivative work to first execute a license for \PDAC\ with INGV in accordance with the terms and conditions of this Agreement. Any derivative work should be clearly marked and renamed to notify users that it is a modiﬁed version and not the original \PDAC\ code distributed by INGV. 

\item Disclaimer of Warranty

Except as expressly set forth in this Agreement, THIS SOFTWARE IS PROVIDED “AS IS” AND INGV MAKES NO REPRESENTATIONS AND EXTENDS NO WARRANTIES OF ANY KIND, EITHER EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO WARRANTIES OR MERCHANTABILITY OR FITNESS FOR A PARTICULAR PURPOSE, OR THAT THE USE OF THE SOFTWARE WILL NOT INFRINGE ANY PATENT, TRADEMARK, OR OTHER RIGHTS. LICENSEE ASSUMES THE ENTIRE RISK AS TO THE RESULTS AND PERFORMANCE OF THE SOFTWARE AND/OR ASSOCIATED MATERIALS. LICENSEE AGREES THAT INGV SHALL NOT BE HELD LIABLE FOR ANY DIRECT, INDIRECT, CONSEQUENTIAL, OR INCIDENTAL DAMAGES WITH RESPECT TO ANY CLAIM BY LICENSEE OR ANY THIRD PARTY ON ACCOUNT OF OR ARISING FROM THIS AGREEMENT OR USE OF THE SOFTWARE AND/OR ASSOCIATED MATERIALS. 

\item Citation

The user agrees that any reports or published results obtained with the Software will acknowledge its use by the appropriate citation as follows: 

\begin{quote}
Neri, A., Esposti Ongaro, T., Macedonio, G., and Gidaspow, D., 2003. Multiparticle simulation of collapsing volcanic columns and pyroclastic flows. J. Geophys. Res., 108, B4, 2202, doi:10.1029/2001JB000508
\end{quote}

\begin{quote}
Esposti Ongaro, T., Cavazzoni, C., Erbacci, G., Neri, A., and Salvetti, M.V., 2007. A parallel multiphase flow code for the 3D simulation of explosive volcanic eruptions. Parallel Computing, doi:10.1016/j.parco.2007.04.003.
\end{quote}

One copy of each publication or report will be supplied to INGV at the addresses listed below in Contact Information.

Electronic documents will include a direct link to the official page: {\tt http://vmsg.pi.ingv.it}

\item Commercial use

Should Licensee wish to make commercial use of the Software, Licensee will contact INGV at the contact infromation addresses listed below to negotiate an appropriate license for such use. Commercial use includes: (1) integration of all or part of the Software into a product for sale, lease or license by or on behalf of Licensee to third parties, or (2) distribution of the Software to third parties that need it to commercialize product sold or licensed by or on behalf of Licensee. 

\item Development

\PDAC\ is being distributed as a research and teaching tool and as such, the PDAC Team encourages contributions from users of the code that might, at INGV’ sole discretion, be used or incorporated to make the basic operating framework of the Software a more stable, ﬂexible, and useful product. Licensees that wish to contribute their code to become an internal portion of the Software may contact the PDAC Team at the addresses listed below in Contact Information.

\end{enumerate}
\newpage
\subsubsection*{Contact Information}

The best contact path for licensing issues is by e-mail to \PDACADDRESS\ or send correspondence to:
\begin{verse}
                             Tomaso Esposti Ongaro\\
                             Istituto Nazionale di Geofisica e Vulcanologia\\
			     Sezione di Pisa\\
                             32, Via della Faggiola\\
			     56126, Pisa - Italy\\
                             FAX: +39-0508311942\\
                             email: ongaro@pi.ingv.it
\end{verse}
\begin{verse}
                             Augusto Neri\\
                             Istituto Nazionale di Geofisica e Vulcanologia\\
			     Sezione di Pisa\\
                             32, Via della Faggiola\\
			     56126, Pisa - Italy\\
                             FAX: +39-0508311942\\
                             email: neri@pi.ingv.it
\end{verse}


